---
title: "Analysis of COVID Vaccine Distribution in Pennsylvania"
subtitle: ""
author: "Team Blue <br> Vinh Hau, Iris Lee, Jason Yan"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(showtext)
library(rvest)
library(janitor)
```


```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, echo = FALSE}
  janssen <- read.csv("../data/janssen.csv")
  pfizer <- read.csv("../data/pfizer.csv")
  moderna <- read.csv("../data/moderna.csv")
```

```{r presentation_theme, echo = FALSE}
style_xaringan(
  title_slide_background_image = "img/title.png",
  title_slide_text_color = "#FFFFFF",
  background_color = "#FFF",
  text_color = "#000080",
  header_color = "#800000",
  inverse_background_color = "#23395D",
  inverse_text_color = "#C0C0C0",
  inverse_header_color = "#FFF",
)
```


class: center, middle

## Are vaccines being distributed fairly within the U.S.?

---
# The Racial Disparities Behind Who Has Received Vaccines

```{r embed_url, echo = FALSE}

include_url("https://www.kff.org/coronavirus-covid-19/issue-brief/latest-data-on-covid-19-vaccinations-race-ethnicity/", height = "600px")

```

---

## The Plan

- First find data on population by race by state

- Then use linear regression to model the number of vaccines based off of the population of each racial group 

- Finally evaluate the accuracy of our model

---

class: inverse, center, middle

# Section 1: Data

---

# Initial Data Sets

- Vaccination Data received from Centers for Disease Control and Prevention (CDC)

- Data on vaccines allocation came in 3 different data sets for the 3 types of vaccines that are largely recommended and authorized in the U.S. (as of the week of April 5th 2021).

---

# janssen.csv, moderna.csv, and pfizer.csv

```{r glimpse_df, echo = FALSE}
glimpse(janssen)

glimpse(moderna)

glimpse(pfizer)
```

---

# Renaming Columns

```{r rename_column}
janssen <- janssen %>%
  rename(
    first_dose = "X1st.Dose.Allocations",
    # second_dose = "2nd Dose Allocations",
    week = "Week.of.Allocations",
    jurisdiction = "Jurisdiction"
  )
moderna <- moderna %>%
  rename(
    first_dose = "X1st.Dose.Allocations",
    second_dose = "X2nd.Dose.Allocations",
    week = "Week.of.Allocations",
    jurisdiction = "Jurisdiction"
  )
pfizer <- pfizer %>%
  rename(
    first_dose = "X1st.Dose.Allocations",
    second_dose = "X2nd.Dose.Allocations",
    week = "Week.of.Allocations",
    jurisdiction = "Jurisdiction"
  )
```

---

# Mutating the data


```{r mutate_df, eval = FALSE}
# First note that the janssen vaccine only has 1 dose. We can fix this with adding a variable `second_dose` with only values of 0. 
janssen_second_dose <- janssen %>%
  mutate(second_dose = 0)
glimpse(janssen_second_dose)


# Then, we add a column called `vaccine` to differentiate the types of vaccines.

janssen_second_dose <- janssen_second_dose %>%
  mutate(vaccine = "janssen")

pfizer <- pfizer %>%
  mutate(vaccine = "pfizer")

moderna <- moderna %>%
  mutate(vaccine = "moderna")

```


```{r mutate_df1, include=FALSE}
# First note that the janssen vaccine only has 1 dose. We can fix this with adding a variable `second_dose` with only values of 0. 
janssen_second_dose <- janssen %>%
  mutate(second_dose = 0)
glimpse(janssen_second_dose)


# Then, we add a column called `vaccine` to differentiate the types of vaccines.

janssen_second_dose <- janssen_second_dose %>%
  mutate(vaccine = "janssen")

pfizer <- pfizer %>%
  mutate(vaccine = "pfizer")

moderna <- moderna %>%
  mutate(vaccine = "moderna")

```

---

# Combining and Writing Data Set

- Then, we combine the datasets into `vaccines` with `rbind`
```{r glimpse_vaccines}
vaccines <- rbind(janssen_second_dose, moderna, pfizer)
glimpse(vaccines)
```

- Finally, we write out the dataset
```{r write_vaccines}
write_csv(vaccines, "../data/vaccines.csv")
```

---

class: inverse, center, middle

# Merging Datasets


---

class: inverse, center, middle

## *demographic.csv*

```{r read_demographic, include =FALSE}

demographic <- read.csv("../data/demographic.csv")

```

---
# demographic.csv
```{r glimpse_demographic}
glimpse(demographic)
```

---

# More tidying

```{r tidy_vaccines_demographic}
# select main variables to be joined, convert case, and rename variables to match snake_case
demographic <- demographic %>%
  select(Location, White, Black, Asian, Hispanic) %>%
  janitor::clean_names() %>%
  rename(
    white_prop = white,
    black_prop = black,
    asian_prop = asian,
    hispanic_prop = hispanic
  )

# change variables name to match

vaccines <- vaccines %>%
  rename(location = "jurisdiction")
```

---

# Joining `vaccine` and `demographic`

```{r vaccine_join_demographic, warning=FALSE}

vaccine_demographic <- left_join(vaccines, demographic, by = "location") %>%
#remove all locations where the demographic data frame does not have, thereby unsuitable for our model
  na.omit()
vaccine_demographic$asian_prop <- lapply(vaccine_demographic$asian_prop, function(x) as.numeric(as.character(x)))
vaccine_demographic <- vaccine_demographic %>%
  mutate(asian_prop = if_else(is.na(asian_prop), 0, as.numeric(asian_prop)))
vaccine_demographic
```

---

class: inverse, center, middle

## *population.csv*

```{r read_population, include =FALSE}

population <- read.csv("../data/population.csv")

```

---
# population.csv
```{r glimpse_population}
glimpse(population)
```

---



# `vaccine_pop`

We essentially repeated the steps above for the population dataframe, but with a little bit more mutating...

```{r read_vaccine_pop, include = FALSE}
vaccine_pop <- read_csv("../data/vaccine_pop.csv")
```

```{r vaccine_pop}
vaccine_pop <- vaccine_pop %>%
  mutate(week = as.Date(week, "%m/%d/%y"))
vaccine_pop
```

---

class: inverse, middle, center

# Section 2: Model


---

# Splitting the Data

First we split the data into train and test data sets.

```{r split_data}
set.seed(1234)
library(tidymodels)
vaccine_split <- initial_split(vaccine_pop)
vaccine_train <- training(vaccine_split)
vaccine_test <- testing(vaccine_split)
dim(vaccine_train)
dim(vaccine_test)
```


---

# Specifying Model

Then we build the model and recipe. We use these to build the workflow. 

```{r model_spec}

vaccine_mod <- linear_reg() %>%
  set_engine("lm")
vaccine_recipe <- recipe(first_dose ~ week + vaccine + white + black + asian + hispanic,
                         data = vaccine_train) %>%
  step_zv(all_predictors())

vaccine_wflow <- workflow() %>%
  add_model(vaccine_mod) %>%
  add_recipe(vaccine_recipe)

```


---

# Fitting Model

We fit the model on the training data.

```{r fit_to_model}
vaccine_fit <- vaccine_wflow %>%
  fit(data = vaccine_train)
vaccine_fit %>%
  tidy()
```

---

# Key Takeaway

On average, one more white person resulted 0.549 more vaccines, while one more black, asian, or hispanic person resulted in around 0.1 more vaccines. 

```{r tidy, echo = FALSE}
vaccine_fit %>%
  tidy()
```



---

class: inverse, middle, center

# Section 3: Evaluate Model


---

# Evaluate Model


```{r predict_vaccine_train}
vaccine_train_pred <- predict(vaccine_fit, vaccine_train) %>%
  bind_cols(data = vaccine_train %>% select(first_dose, location))
vaccine_train_pred
```

---

# Residual Graph
```{r residual, echo=FALSE}
vaccine_mod_fitted <- vaccine_mod %>% 
  fit(first_dose ~ week + vaccine + white + black + asian + hispanic, data = vaccine_train)
vaccine_aug <- augment(vaccine_mod_fitted$fit)

vaccine_aug %>%
  ggplot(aes(x=.fitted, y=.resid)) +
  geom_point()
```


---

# R^2
```{r rmse_rsq_train}
rmse(vaccine_train_pred, truth = first_dose, estimate = .pred)
rsq(vaccine_train_pred, truth = first_dose, estimate = .pred)
```

---


```{r predict_vaccine_test}
vaccine_test_pred <- predict(vaccine_fit, vaccine_test) %>%
  bind_cols(data = vaccine_test %>% select(first_dose, location))
vaccine_test_pred
```

---

# R^2
```{r rmse_rsq_test}
rmse(vaccine_test_pred, truth = first_dose, estimate = .pred)
rsq(vaccine_test_pred, truth = first_dose, estimate = .pred)



```

---

# Conclusion

The residual graph displays a random distribution meaning that a linear regression model is the correct model for our data. Further our RMSE is small and our R^2 is large indicating that our model quite accurately predicts the number of vaccinations based on racial distribution. 

---

# Limitations

- Implicit bias
- Correlation vs. Causation
- Third confounding variable
- Overgeneralization

---

# Recommendations

- Establish cause
- Find more focused datasets


